Merging merges, more or less

### 2015-03-17 2015mergingmerges/section/02-Machines.tex
add Close state to machines

         \> $~|$\> @Release@  \>     \> $n$         \\
         \>     \> (Read inputs must be released) \\

@@

Each of these state types requires a corresponding set of output transitions to be defined.
\begin{tabbing}
MMMMMMMM \= M \= \kill
$\gamma(@Pull @n)$
                            \> $=$ \> \{@Some@ n, @None@\} \\
$\gamma(@Release @n)$
                            \> $=$ \> \{@Unit@\} \\
$\gamma(@Out @   f~n)$
                            \> $=$ \> \{@Unit@\} \\
$\gamma(@OutDone @ n)$
                            \> $=$ \> \{@Unit@\} \\
$\gamma(@If @    f~n)$
                            \> $=$ \> \{@True@, @False@\} \\
$\gamma(@Update @f~n)$
                            \> $=$ \> \{@Unit@\} \\
$\gamma(@Skip@      )$
                            \> $=$ \> \{@Unit@\} \\
$\gamma(@Done@      )$
                            \> $=$ \> \{\} \\
\end{tabbing}

@@

>>>>

         \> $~|$\> @Release@  \>     \> $n$         \\
         \>     \> (Values must be released before the input can be pulled again) \\

         \> $~|$\> @Close@    \>     \> $n$         \\
         \>     \> (A named input may still have data, but we will not read it) \\

@@

Each of these state types requires a corresponding set of output transitions to be defined.
\begin{tabbing}
MMMMMMMM \= M \= \kill
$\gamma(@Pull @n)$
                            \> $=$ \> \{@Some@ n, @None@\} \\
$\gamma(@Release @n)$
                            \> $=$ \> \{@Unit@\} \\
$\gamma(@Close @n)$
                            \> $=$ \> \{@Unit@\} \\
$\gamma(@Out @   f~n)$
                            \> $=$ \> \{@Unit@\} \\
$\gamma(@OutDone @ n)$
                            \> $=$ \> \{@Unit@\} \\
$\gamma(@If @    f~n)$
                            \> $=$ \> \{@True@, @False@\} \\
$\gamma(@Update @f~n)$
                            \> $=$ \> \{@Unit@\} \\
$\gamma(@Skip@      )$
                            \> $=$ \> \{@Unit@\} \\
$\gamma(@Done@      )$
                            \> $=$ \> \{\} \\
\end{tabbing}

@@

\subsection{Invariants}
Not all machines are valid or meaningful, and we wish to rule them out.
For example, a machine with an initial state of $@If@~(x>5)~n$ is meaningless because there is no local $x$ that has been read from an input. 
On the other hand, some programs could be meaningful, but requiring them to be in a particular form simplifies fusion, later.
An example of such is a program that does not release its input before pulling again; a sympathetic code extraction would be to simply release before every pull, but the explicit pulls makes synchronising two machines easier.

The invariants are:
\begin{itemize}
\item each function mentioned in @Out@, @If@ or @Update@ can only refer to previously read values;
\item each successful @Pull@ must be @Released@ before another @Pull@ can be made;
\item all inputs must either be finished by an empty @Pull@ or @Close@d at @Done@;
\item all outputs must be finished with @OutDone@ at @Done@;
\item inputs cannot be read after an empty @Pull@ or after they are @Close@d;
\item outputs cannot be written after they are finished with @OutDone@.
\end{itemize}

\begin{code}
check_machine m
 = check_state (_init m) (M.singleton (_init m) S.empty)
 where
  check_state l acc
   | Just t <- M.lookup l (_trans m)
   , Just s <- M.lookup l acc
   = case t of
      Pull n l1 l2
       | not $ Value    n `S.member` s
       , not $ Finished n `S.member` s
       ->  go l1 (S.insert (Value n) s) acc
       >>= go l2 (S.insert (Finished n) s)

      Release n l'
       | Value n `S.member` s
       -> go l' (S.delete (Value n) s)
          acc

      Close   n l'
       | not $ Value n `S.member` s
       -> go l' (S.insert (Finished n) s)
          acc

      Update f l'
       | all (`S.member` s) (fvs f)
       -> go l' s acc

      If f l1 l2
       | all (`S.member` s) (fvs f)
       -> go l1 s acc >>= go l2 s

      Out f l'
       | all (`S.member` s) (fvs f)
       -> go l' (S.insert (Value (_state f)) s) acc

      OutFinished n l'
       | not $ Finished n `S.member` s
       -> go l' (S.insert (Finished n) s) acc
       | otherwise

      Skip l'
       -> go l' s acc

      Done
       | all (\n -> Finished n `S.member` s) 
       $ S.toList
       $ S.union inputs outputs
       -> Right acc

  go l s m
   | Just s' <- M.lookup l m
   = let sI  = S.intersection s s'
         sD  = S.difference   s s'
     -- Allow 
     in  if         s == s'
         then       Right m
         else if    all is_output (S.toList sD)
         then       check_state l (M.insert l sI m)
         else       Error
   | otherwise
   = check_state l (M.insert l s m)
\end{code}

### 2015-03-17 2015mergingmerges/section/03-Access.tex
add Close state to machines

If the first machine is producing an output on channel $n$ and the second machine uses $n$, the first machine would usually have to wait until the second machine is ready to accept new data.
However, if the second machine is @Done@, the first machine may output to the channel as frequently as it likes.
Note that the output channel $n$ may have multiple consumers, so continuing to output after the second machine consumer has finished is not as futile as it may seem at first.
\TODO{Although, all the combinators so far use the whole input, and anyway checking for @Done@ of machine is too fine-grained if machine already has multiple combinators; perhaps an ``I promise not to read from X any more'' state would be interesting.}

>>>>

If the first machine is producing an output on channel $n$ and the second machine uses $n$, the first machine would usually have to wait until the second machine is ready to accept new data.
However, if the second machine is @Done@, the first machine may output to the channel as frequently as it likes.
Note that the output channel $n$ may have multiple consumers, so continuing to output after the second machine consumer has finished is not as futile as it may seem at first.
\TODO{Although, all the combinators so far use the whole input, and anyway checking for @Done@ of machine is too fine-grained if machine already has multiple combinators; perhaps an ``I promise not to read from X any more'' state would be interesting. Actually, @zip@ doesn't consume the whole input.}
